---
title: "Final Project Markdown"
author: "ian johnson"
date: "May 1, 2016"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(plyr)
library(lattice)
library(lubridate)
library(plyr)
library(lattice)
library(lubridate)
library(Hmisc)
library("latticeExtra")
library(lattice)
library(ggplot2)
library(stats)
library(graphics)

setwd('~/Desktop/Columbia/EDAV/Final Project')
df = read.csv('data/2015.csv')
finance = mdb.get('AVROLL.mdb')
```



## Abstract

Our analysis of the 311 complaints data was focused on determining the factors that played into variance in reponse time for different complaint types. We were interested in which of these factors might play the largest role in causing an increase in response time, so that we might be able to draw conclusions on how best to remedy handle the negative effects of this variable. To investigate response time we chose the top 7 complaints by count, and looked into several independent variables from the original dataset, as well as from other datasets from alternative sources. We then form a predictive model that takes into account these various factors, and allows us to predetermine a response time given certain initial conditions for the complaint. 


## Data Cleanup

There were several steps that we took to massage the data into a more useful form in order to perform our analyses. First off, we chose to use only data from the year 2015, in order to make sure that our analyses were being performed on a consistend subset of the data, and to give us faster times for further subsetting and data manipulation. 


```{r cleanup}

###########################
# Complaints data section #
###########################

# Remove bad zipcodes
df = df[!is.na(df$Incident.Zip),]
df$Incident.Zip = substr(df$Incident.Zip, 0, 5)

# Subset to top seven complaint types
df = df[ which(
    df$Complaint.Type == "HEAT/HOT WATER" |
    df$Complaint.Type == "Blocked Driveway" |
    df$Complaint.Type == "Illegal Parking" |
    df$Complaint.Type == "UNSANITARY CONDITION" |
    df$Complaint.Type == "PAINT/PLASTER" |
    df$Complaint.Type == "PLUMBING" |
    df$Complaint.Type == "Noise - Street/Sidewalk"
  ), ]

# Column for response time
closed_dates = mdy_hms(df$Closed.Date)
created_dates = mdy_hms(df$Created.Date)
df$response_time = difftime(closed_dates, created_dates, units="hours")
df$response_time = as.numeric(df$response_time, units="hours")

# Slimmed down version of relevant columns
df_slim = df[,c("Created.Date","Agency","Complaint.Type", "Incident.Zip", "response_time")]
```








```{r financial_prediction}

avroll = finance$avroll
condensed = finance$`Condensed Roll Description`

# Calculate mean building price per zipcode
avg_vals = sapply(split(avroll$FULLVAL, avroll$ZIP), mean) 
avg_vals_df = data.frame(zipcode=names(avg_vals), value=avg_vals, row.names=NULL)
write.csv(avg_vals_df, file='zipcode_average_values.csv')


################################
# Building prices data section #
################################

# Calculate mean building price per zipcode
avg_prices = sapply(split(avroll$FULLVAL, avroll$ZIP), mean) 
avg_prices_df = data.frame(zipcode=names(avg_prices), value=avg_prices, row.names=NULL)


########################
# Combine into Trellis #
########################
df_merged = merge(df, avg_prices_df, "left", by.x="Incident.Zip", by.y="zipcode")

# Aggregating
rel_df = data.frame(zip=df_merged$Incident.Zip,
                    type=df_merged$Complaint.Type,
                    time=df_merged$response_time,
                    borough=df_merged$Borough)
rel_df$zip = as.factor(rel_df$zip)

rel_df_avg = aggregate(time ~ zip + type, rel_df, mean)
# Merge back in the prices and neighborhoods after the aggregation removes them
rel_df_avg_with_prices = merge(rel_df_avg, avg_prices_df, by.x="zip", by.y="zipcode")

zips_boroughs = unique(df[c("Incident.Zip", "Borough")])

rel_df_avg_with_prices_boroughs = merge(rel_df_avg_with_prices,
                                        zips_boroughs,
                                        all.x=TRUE,
                                        all.y=FALSE,
                                        by.x="zip",
                                        by.y="Incident.Zip")

pg = ggplot(rel_df_avg_with_prices_boroughs, aes(value, time)) + 
  geom_point(shape = 1, aes(color = Borough)) + 
  
  scale_y_log10(name="Average Response Time",
                breaks=c(1, 10, 120, 1460, 8760),
                labels=c("1 Hour", "10 Hours", "5 Days", "2 Months", "1 Year")) + 
  
  scale_x_log10(name="Average Building Value",
                breaks=c(1000000, 10000000, 100000000),
                labels=c("1M", "10M", "100M")) + 
  
  facet_wrap(~type,nrow=2) + 
  ggtitle("Average Building Value vs Average Response Time per Zipcode across Top 7 Complaints")

pg + theme(axis.text=element_text(size=12),
        axis.title=element_text(size=16)
        # plot.margin=unit(c(10,10,10,10),"mm")
        )

print(pg)

# -> No relationship b/w avg building price and response time. Show lm stats to confirm:
fit = lm(value ~ time, data = rel_df_avg_with_prices)
summary(fit)
plot(value ~ time, data = rel_df_avg_with_prices)
abline(fit)

```



